---
title: "p8105_hw3_hx2263"
author: "Tiffany Xi"
date: "10/5/2018"
output: github_document
---
(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)(commenting on the results of each)
```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

library(ggplot2)
```

# Problem 1

* Load the `BRFSS` data from the p8105.datasets package.

```{r df1_import, message = FALSE}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

* Data cleaning

```{r df1_clean}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
#select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
  filter(response == "Excellent"| response == "Very good"| response == "Good"| response == "Fair"| response == "Poor") %>% 
  mutate(response = as.factor(response))
```

### Answers

1.In 2002, which states were observed at 7 locations?

```{r}
brfss %>% 
  filter(year == 2002) %>% 
  distinct(locationdesc, .keep_all = TRUE) %>% 
  count(locationabbr) %>% 
  filter(n == 7)
```

In 2002, the states which were observed at 7 locations are `r (brfss %>% filter(year == 2002) %>% distinct(locationdesc, .keep_all = TRUE) %>% count(locationabbr) %>% filter(n == 7))$locationabbr`.

2.Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r spaghetti_plot}
brfss %>% 
  distinct(year, locationdesc, .keep_all = TRUE) %>% 
  filter(year > 2001 & year < 2011) %>% 
  group_by(year, locationabbr) %>% 
  summarize(number = n()) %>% 
  ggplot(aes(x = year, y = number, color = locationabbr)) +
  geom_point(size = .5, alpha = .5) +
  geom_line(alpha = .6) +
  scale_x_continuous(breaks = c(2002, 2003, 2004, 2005, 2006, 
                                2007, 2008, 2009, 2010), 
                     labels = c("2002", "2003", "2004", "2005", "2006", 
                                "2007", "2008", "2009", "2010")) +
  theme_bw()
```

3.Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r mean_sd_table}
brfss %>% 
  filter(locationabbr == "NY" & (year == 2002 | year == 2006 | year == 2010) & response == "Excellent") %>% 
  group_by(year) %>% 
  summarize(mean_excellent_prop = mean(data_value, na.rm = TRUE),
            sd_excellent_prop = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(digits = 1)
```

4.For each year and state, compute the average proportion in each response category (taking the average across locations in a state). 

```{r avg_prop}
avg_prop_rep = brfss %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(avg_prop = round(mean(data_value, na.rm = TRUE), digits = 2))
```

Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r five_panel}
ggplot(data = avg_prop_rep, aes(x = year, y = avg_prop, color = locationabbr)) +
    facet_grid( ~ response) +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom")
```

```{r five_plot}
excel_rep = avg_prop_rep %>% filter(response == "Excellent")
vg_rep = avg_prop_rep %>% filter(response == "Very good")
good_rep = avg_prop_rep %>% filter(response == "Good")
fair_rep = avg_prop_rep %>% filter(response == "Fair")
poor_rep = avg_prop_rep %>% filter(response == "Poor")

panel_plot = function(df){
  ggplot(data = df, aes(x = year, y = avg_prop, color = locationabbr)) +
    geom_line() +
    theme(legend.position = "bottom")
}

panel_plot(excel_rep)
panel_plot(vg_rep)
panel_plot(good_rep)
panel_plot(fair_rep)
panel_plot(poor_rep)  
  
```

```{r 什么鬼}
ggplot(data = avg_prop_rep, aes(x = year, y = avg_prop, color = locationabbr)) +
    facet_grid( ~ response) +
    geom_point(alpha = .5) +
    geom_smooth(se = FALSE) +
    theme_bw() +
    theme(legend.position = "bottom")

```

# Problem 2

* Load the `instacart` data from the p8105.datasets package and take a look.

```{r}
instacart = p8105.datasets::instacart
head(instacart)
```

### Description

The `instacart` dataset has `r ncol(instacart)` columns and `r nrow(instacart)` rows, which means it contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each row in the dataset is a product information from an order. The structure of this dataset is *Tabular Data*. As far as I am concerned, this dataset describes the orders and reorder detail from customer `user_id` though Instacart app.

The key variables are `order_id`, `product_id` and `product_name`. We can see the user `user_id` bought `product_id` which is `product_name` belonging to `department` and located in `aisle`. We can also know how frequent one customer use this app to do grocery shopping according to `days_since_prior_order` and `order_number`. 

For example, the `user_id`1 made `order_id`1187899. He/she bought 11 items in `order_number`11 `days_since_prior_order`14 days after his/her last shopping, among which the `add_to_cart_order`9 was the first time order, other 10 items were reorders. `product_id`196 is Soda in `aisle`soft drinks `aisle_id`77 in `department`beverages which id is 7.

### Answers

1.How many aisles are there, and which aisles are the most items ordered from?

There are `r nrow(instacart %>% count(aisle))` aisles.

`r (instacart %>% count(aisle) %>% mutate(aisle_ranking = min_rank(desc(n))) %>% filter(aisle_ranking < 4))$aisle` aisle are the top 3 aisles which most items ordered from.
Among those, `r (instacart %>% count(aisle) %>% mutate(aisle_ranking = min_rank(desc(n))) %>% filter(aisle_ranking < 2))$aisle` aisle ranks first.


2.Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
ggplot(data = instacart, aes(x = aisle)) +
  geom_bar(position = position_stack(reverse = TRUE)) +
  coord_flip() +
  labs(
      title = "The Number of Items Ordered in Each Aisle",
      y = "the Number of Items (counts)",
      x = "Aisle",
      caption = "Data from the p8105.package"
    ) +
  theme_bw()
```


3.Make a table showing the most popular item aisles "baking ingredients", "dog food care", and "packaged vegetables fruits"

```{r}
instacart %>% 
  filter(aisle == "baking ingredients"| aisle == "dog food care"| aisle == "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(item_ranking = min_rank(desc(n))) %>% 
  filter(item_ranking < 2) %>% 
  knitr::kable()
```

4.Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples"| product_name == "Coffee Ice Cream" ) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour) %>% 
  knitr::kable(digits = 2)
```


# Problem 3

* Load the `NY NOAA` data from the p8105.datasets package
```{r}
noaa = p8105.datasets::ny_noaa
```

### Description
The `noaa` dataset has `r ncol(noaa)` columns and `r nrow(noaa)` rows, which means it contains `r nrow(noaa)` observations and `r ncol(noaa)` variables. The structure of this dataset is *Tabular Data*. 

The key variables are `id`, `date`, `prcp`, `snow`, `snwd`, `tmax` and `tmin`. Each row in the dataset is a weather information including precipitation`prcp`, snowfall`snow`, snow depth`snwd`, maximum temperature`tmax` and min temperature`tmin` from one single weather station`id` in one day`date`. This dataset describes five core variables for all New York state weather stations from January 1, 1981 through December 31, 2010 according to `r nrow(noaa %>% distinct(id))` weather stations.

Missing values is a huge issue in this dataset, the number of missing number in five core variables are `r sum(is.na(noaa$prcp))`, `r sum(is.na(noaa$snow))`, `r sum(is.na(noaa$snwd))`, `r sum(is.na(noaa$tmax))` and `r sum(is.na(noaa$tmin))`. Those are huge numbers and almost half of the whole rows doesn't have tmax and tmin. 

### Answers

1.Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

```{r}
noaa_tidy = noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(pcrp_mm = ifelse(is.na(prcp), NA, prcp / 10),
         snow_mm = snow,
         tmax_ºC = ifelse(is.na(tmax), NA, as.integer(tmax)/ 10),
         tmin_ºC = ifelse(is.na(tmin), NA, as.integer(tmin)/ 10)) %>% 
  select(-prcp, -snow, -tmax, -tmin)
```

For snowfall, what are the most commonly observed values? Why?

```{r}
noaa_tidy %>% 
  count(snow_mm) %>% 
  mutate(snow_ranking = min_rank(desc(n))) %>% 
  filter(snow_ranking < 2)
```

* For snowfall, 0 are the most commonly observed values, because snow only happens in winter which is one over four seasons, so other seasons take up three fourth days. Also, only when temperature goes below zero can snow, which takes up a smaller percentage of New York State winter temp. So 0mm snowfall is the most commonly observed values.


2.Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
noaa_tidy %>% 
  filter(month == "01" | month == "07") %>% 
  group_by(year, month) %>% 
  summarize(mean_tmax = mean(tmax_ºC, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax, color = month)) +
  geom_point(size = 4)

noaa_tidy %>% 
  filter(month == "01" | month == "07") %>% 
  group_by(year, month) %>% 
  summarize(mean_tmin = mean(tmin_ºC, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmin, color = month)) +
  geom_point(size = 4)
```

3.Make a two-panel plot showing 
(i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option);
(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
ggplot(data = noaa_tidy, aes(x = year, y = tmax_ºC)) +
  geom_point()
```

```{r}
snow_year = noaa_tidy %>% 
  filter(snow_mm > 0 & snow_mm < 100) %>% 
  group_by(year, snow_mm)

#ggplot(data = snow_year, aes(x = mean(snow_mm, na.rm = TRUE))) +
  #geom_histogram()
```

