---
title: "p8105_hw3_hx2263"
author: "Tiffany Xi"
date: "10/5/2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

library(ggplot2)
```

#Problem 1

Load the `BRFSS` data from the p8105.datasets package

```{r data_import, message = FALSE}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

Data cleaning

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  #select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>% 
  filter(response == "Excellent"| response == "Very good"| response == "Good"| response == "Fair"| response == "Poor") %>% 
  mutate(response = as.factor(response))
```

Using this dataset, answer the question

1.In 2002, which states were observed at 7 locations?

```{r}
brfss %>% 
  filter(year == 2002) %>% 
  distinct(locationdesc, .keep_all = TRUE) %>% 
  count(locationabbr) %>% 
  filter(n == 7)
```

In 2002, the states which were observed at 7 locations are `r (brfss %>% filter(year == 2002) %>% distinct(locationdesc, .keep_all = TRUE) %>% count(locationabbr) %>% filter(n == 7))$locationabbr`.

2.Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}
brfss %>% 
  distinct(year, locationdesc, .keep_all = TRUE) %>% 
  filter(year > 2001 & year < 2011) %>% 
  group_by(year, locationabbr) %>% 
  summarize(number = n()) %>% 
  ggplot(aes(x = year, y = number, color = locationabbr)) +
  geom_point(size = .5, alpha = .5) +
  geom_line(alpha = .6) +
  scale_x_continuous(breaks = c(2002, 2003, 2004, 2005, 2006, 
                                2007, 2008, 2009, 2010), 
                     labels = c("2002", "2003", "2004", "2005", "2006", 
                                "2007", "2008", "2009", "2010")) +
  theme_bw()
```

3.Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss %>% 
  filter(locationabbr == "NY" & (year == 2002 | year == 2006 | year == 2010) & response == "Excellent") %>% 
  group_by(year) %>% 
  summarize(mean_excellent_prop = mean(data_value, na.rm = TRUE),
            sd_excellent_prop = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(digits = 1)
```

4.For each year and state, compute the average proportion in each response category (taking the average across locations in a state). 

```{r}
avg_prop_rep = brfss %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(avg_prop = round(mean(data_value, na.rm = TRUE), digits = 2))
```

Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
ggplot(data = avg_prop_rep, aes(x = year, y = avg_prop, color = locationabbr)) +
    facet_grid( ~ response) +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom")
```

```{r}
excel_rep = avg_prop_rep %>% filter(response == "Excellent")
vg_rep = avg_prop_rep %>% filter(response == "Very good")
good_rep = avg_prop_rep %>% filter(response == "Good")
fair_rep = avg_prop_rep %>% filter(response == "Fair")
poor_rep = avg_prop_rep %>% filter(response == "Poor")

panel_plot = function(df){
  ggplot(data = df, aes(x = year, y = avg_prop, color = locationabbr)) +
    geom_line() +
    theme(legend.position = "bottom")
}

panel_plot(excel_rep)
panel_plot(vg_rep)
panel_plot(good_rep)
panel_plot(fair_rep)
panel_plot(poor_rep)  
  
```

```{r 什么鬼}
ggplot(data = avg_prop_rep, aes(x = year, y = avg_prop, color = locationabbr)) +
    facet_grid( ~ response) +
    geom_point(alpha = .5) +
    geom_smooth(se = FALSE) +
    theme_bw() +
    theme(legend.position = "bottom")

```

